Competing Bandits: The Perils of Exploration under Competition

### 2018-09-05 paper/content/introduction.tex
Edits in inverted-U and "barriers" paragraphs.

We consider varying levels of competition, from permanent monopoly to \emph{temporary monopoly} (a duopoly with an early entrant), to a duopoly in which both firms start at the same time. We find that a ``greedy algorithm" that does not explicitly explore is most beneficial under duopoly. This algorithm also prevails under monopoly, simply because it tends to be easier to deploy. Whereas a temporary monopoly incentivizes more advanced exploration algorithms that perform better in the long run. The disincentives to explore under duopoly arise entirely because of ``reputational costs", rather than R\&D costs (which are absent from our model).

@@

Our findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy. \gadelete{In this scenario, the incumbent firm enjoys both a ``data advantage" and a ``reputational advantage". We further investigate which of the two is a stronger barrier to entry. We find that both are strong barriers on its own: removing either one substantially helps the late entrant. Further, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm.}

\gaedit{Being first in the market gives firms free data to learn from (a ``data advantage") as well a larger reputation compared to an entrant (a ``reputation advantage"). We investigate which of the two is a stronger barrier to entry. We find that both are strong barriers on their own: removing either one still gives the incumbent a substantially larger share of the market compared to the case where both are removed. Further, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm. This occurs because the lack of initial competition allows the incumbent to gather data on actions that may be harder to gather in competition (due to reputational consequences) but that are useful for learning and maximizing long-run reputation.
}

>>>>

\asedit{We consider a \emph{permanent duopoly} in which both firms start at the same time, as well as \emph{temporary monopoly}: a duopoly with an early entrant. Accordingly, the intensity of competition in the model varies from ``permanent monopoly" (just one firm) to ``early entry" (in the temporary monopoly) to permanent duopoly to ``late entry" (temporary monopoly from the incumbent's perspective). We find that a ``greedy algorithm" that does not explicitly explore is most beneficial under duopoly, and even more so under ``late entry".} This algorithm also prevails under monopoly, simply because it tends to be easier to deploy. Whereas a temporary monopoly incentivizes more advanced exploration algorithms that perform better in the long run. The disincentives to explore under duopoly arise entirely because of ``reputational costs", rather than R\&D costs (which are absent from our model, aside from the tie-breaking).

@@

Our findings on temporary monopoly shed light on the ``first-mover advantage" phenomenon in the digital economy. Being first in the market gives firms free data to learn from (a ``data advantage") as well as \asedit{a more definite (and possibly better)} reputation compared to an entrant (a ``reputation advantage"). We investigate which of the two is a stronger barrier to entry. We find that both are strong barriers on their own: removing either one still gives the incumbent a substantially larger share of the market compared to the case where both are removed. Further, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm.

\OMIT{This occurs because the lack of initial competition allows the incumbent to gather data on actions that may be harder to gather in competition (due to reputational consequences) but that are useful for learning and maximizing long-run reputation.}

\OMIT{ %%%%%
\gadelete{In this scenario, the incumbent firm enjoys both a ``data advantage" and a ``reputational advantage". We further investigate which of the two is a stronger barrier to entry. We find that both are strong barriers on its own: removing either one substantially helps the late entrant. Further, the data advantage tends to be a stronger barrier when the incumbent commits to a more advanced bandit algorithm.}
} %%%%%%
